Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

We seek to address the problem of evaluating the use cases for a collision,
to determine whether or not a collision affects a particular system.
The central thesis of this research is that logical cryptanalysis
allows us to formalize the structure inherent to the underlying hash
function such that new collisions can be found within reasonable time
on commodity hardware.  We assert that when such structure exists, it
is \textbf{imperitive to discontinue use of such hash functions as
  they are fundamentally untrustworthy for critical applications.}  An
example of such structure showing the unit distance neighborhood
(defined in Section \ref{Sec:Terminology}) of the known collisions
from Dobbertin and Kasselman is presented in Figure \ref{Fig:Graph-Intro}.
This paper contributes several techniques to logical cryptanalysis and uses these
techniques to derive metrics of the trustworthiness for hash functions. We claim
the following results are novel in the field:
\begin{itemize}
    \item We define the neighborhood of a class of collisions and show that
        it is frequently non-empty.
    \item We discuss techniques for finding useful collisions within a
        class of collisions using logical cryptanalysis.
    \item We propose additional techniques for building new classes of
        collisions and show that collisions in MD4 can be found inductively.
    \item We propose five universal metrics for evaluating the utility of
        a class of collisions.
\end{itemize}
We propose that collision classes of high utility are closer to a second
preimage attack in the amount of flexibility they provide to an attacker,
whereas classes of low utility may not affect all systems which use a
particular hash function.

The remainder of this paper is organized as follows. In Section
\ref{Sec:Related} we discuss previous results and show how our research
expands upon prior work. In Section \ref{Sec:Terminology}, we discuss the
terminology and notations we use throughout the remainder of the paper.
In Section \ref{Sec:Intuition}, we give the intuition behind measuring
the utility of a collision and motivate why it is useful. In Section
\ref{Sec:Logical}, we present new techniques in logical cryptanalysis to
analyze collisions in hash functions. In Section \ref{Sec:Empirical}, we
discuss the utility of existing collisions under the metrics described in
\ref{Sec:Intuition}. In Section \ref{Sec:Collision}, we discuss new
differential paths and the effectiveness of various techniques presented
in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Impacts}, we
discuss broader impacts of this research.

@@

This paper draws heavily upon the public cryptanalysis of MD4, the wide
availability of high quality SAT solvers, and previous work in logical
cryptanalysis. In particular, we look at ways that SAT solving can aid
cryptanalysis instead of using cryptanalysis to benchmark SAT solvers. We
then apply them to aid the understanding of attacks against real world
systems, and propose general metrics that target different types of systems.
Towards this, we look at prior work in two main areas: differential and logical
cryptanalysis.

@@

Our work removes the need for finding message modification techniques by
specifying the differential path as part of the SAT formula and letting the
solver find a pair of satisfying messages which follow the differential path
and produce a collision. Further constraints can then be placed upon this
model, such as a chosen prefix or desired start state. We believe performing
these attacks by hand to be difficult, and note that developed message
modification techniques may affect chosen prefixes. Thus, we seek to develop
techniques to replace differential cryptanalysis with logical cryptanalysis.

@@

However, much of the work in this area is focused on benchmarking SAT solvers,
and not on evaluating techniques for using SAT solvers to aid reasoning and
understanding. E. Homsirikamol et al.\ show that SAT solvers can be used to
brute force finding collisions in SHA-3, but with limited utility and limited
to a small number of rounds \cite{Homsirikamol2012}. On the other hand,
advances by V. Nossum in his master's thesis shows that additional work on
encoding is also important: a new 32-bit modular addition circuit produced
shorter run times \cite{nossum_2012}.

@@

\subsection{On Trustworthy Computing}

\textbf{TODO - Assigned @erozier}

@@

\section{Discussion of Impacts} \label{Sec:Impacts}

@@

\section{Future Work}

>>>>

We seek to address the problem of evaluating the use cases for a collision,
to determine whether or not a collision affects a particular system.
The central thesis of this research is that logical cryptanalysis
allows us to formalize the structure inherent to the underlying hash
function such that new collisions can be found within reasonable time
on commodity hardware.  We assert that when such structure exists, it
is \textbf{imperitive to discontinue use of such hash functions as
  they are fundamentally untrustworthy for critical applications.}  An
example of such structure showing the unit distance neighborhood
(defined in Section \ref{Sec:Terminology}) of the known collisions
from Dobbertin and Kasselman is presented in Figure \ref{Fig:Graph-Intro}.
This paper contributes several techniques to logical cryptanalysis and uses these
techniques to derive metrics of the trustworthiness for hash functions. We claim
the following results are novel in the field:
\begin{itemize}
    \item We define the neighborhood of a class of collisions and show that
        it is frequently non-empty.
    \item We discuss techniques for finding useful collisions within a
        class using logical cryptanalysis.
    \item We propose additional techniques for building new classes of
        collisions and show that classes of collisions in MD4 can be found inductively.
    \item We propose five universal metrics for evaluating the utility of
        a class of collisions.
\end{itemize}
We propose that collision classes of high utility are closer to a second
preimage attack in the amount of flexibility they provide to an attacker,
whereas classes of low utility may not affect all systems which use a
particular hash function.

The remainder of this paper is organized as follows. In Section
\ref{Sec:Related} we discuss previous results and show how our research
expands upon prior work. In Section \ref{Sec:Terminology}, we discuss the
terminology and notations we use throughout the remainder of the paper.
In Section \ref{Sec:Intuition}, we give the intuition behind measuring
the utility of a collision and motivate why it is useful. In Section
\ref{Sec:Logical}, we present new techniques in logical cryptanalysis to
analyze collisions in hash functions. In Section \ref{Sec:Empirical}, we
discuss the utility of existing collisions under the metrics described in
\ref{Sec:Intuition}. In Section \ref{Sec:Collision}, we discuss new
differential paths and the effectiveness of various techniques presented
in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Impacts}, we
discuss the broader impacts and conclusions of this research, and
conclude with Section \ref{Sec:Conclusion} discussing future work and
extensions in Section \ref{Sec:Future}.

@@

This paper draws heavily upon the public cryptanalysis of MD4, the wide
availability of high quality SAT solvers, and previous work in logical
cryptanalysis. In particular, we look at ways that SAT solving can aid
cryptanalysis instead of using cryptanalysis to benchmark SAT
solvers.  A critical assumption of our use of SAT solvers as a means
to discover structures in hash functions which lead to untrustworthy
behavior is the insight that while SAT is, in general, a hard problem
for many specific and practical instances current SAT-solving
technology is quite efficient \cite{vardi2010p}. We
then apply them to aid the understanding of attacks against real world
systems, and propose general metrics that target different types of systems.
Towards this, we look at prior work in two main areas: differential and logical
cryptanalysis.

@@

Our work advances the current state of the art by removing the need for finding message modification techniques by
specifying the differential path as part of the SAT formula and letting the
solver find a pair of satisfying messages which follow the differential path
and produce a collision. Further constraints can then be placed upon this
model, such as a chosen prefix or desired start state. We believe performing
these attacks by hand to be difficult, and note that developed message
modification techniques may affect chosen prefixes. Thus, we seek to develop
techniques to replace differential cryptanalysis with logical cryptanalysis.

@@

Much of the work in this area is focused on the use of these problems
to benchmark the performance of SAT solvers,
and not to evaluate techniques for using SAT solvers to aid reasoning and
understanding. E. Homsirikamol et al.\ show that SAT solvers can be used to
brute force finding collisions in SHA-3, but with limited utility and limited
to a small number of rounds \cite{Homsirikamol2012}. On the other hand,
advances by V. Nossum in his master's thesis shows that additional work on
encoding is also important: a new 32-bit modular addition circuit produced
shorter run times \cite{nossum_2012}.

@@

\subsection{Trustworthy Computing}

Hash functions see wide usage in modern trustworthy computing
applications, where they are used to establish the authenticity and
provenance of data, the credibility of image data
\cite{friedman1993trustworthy}, sensor identity and trust
\cite{wong2006dynamic}, meta-data validity \cite{bayer1993improving},
hardware authorization \cite{sirer2011logical}, and other measures of
trust in system and mission critical systems.  
The use of deprecated hashes for commit verification in
git\cite{loeliger2012version}, and for patch identity and trust
verification \cite{kuhn2005secure}, or segment verification in
peer-to-peer decentralized networks \cite{lua2005survey} is quite
common.  The use of deprecated hashes poses a serious problem, and can
compromise an otherwise trustworthy system or platform.  A key
component missing from the existing literature is a good metric which
can be used to assess the trustworthiness of a hash, and its
suitability for a given use case.

Such a metric needs to be one which allows for a variable which
accounts for difficulty, so that trustworthiness can be assessed in a
meaningful way in an environment with changing hardware capabilities.

@@

\section{Discussion} \label{Sec:Impacts}

@@

\section{Future Work} \label{Sec:Future}
